---
title: "Homework 5"
author: "Sharleen Price spp2122"
date: "4/25/2018"
output: 
  pdf_document: 
    latex_engine: xelatex
---


1) reject-accept
```{r}

#1) function
f<-  function(x){
  return(ifelse((-1<=x | 2>=x),(1/9)*(4-x^2),0))
}

x <- seq(-1, 2, length = 100)
f.max=max(f(x))
f.max
plot(x, f(x), type="l", ylab="f(x)")

#envelope
e <- function(x) {
  return(ifelse((-1 > x | x > 2), Inf, f.max))
}
lines(c(-1, -1), c(f(-1), e(-1)), lty = 1)
lines(c(2, 2), c(0, e(2)), lty = 1)
lines(x, e(x), lty = 1)

#accept-reject algorithm 
alp=0
n.samps <- 1000   # number of samples desired> 
n       <- 0                     # counter for number samples accepted> 
samps   <- numeric(n.samps) # initialize the vector of output> 
while (n < n.samps) 
  {
   y <- runif(1,-1,2)    #random draw from g   
   u <- runif(1) 
   if (u < f(y)/e(y))
     alp=alp+1
     {n <- n + 1     
     samps[n] <- y  } #simulated data
}
alp
#histogram
hist(samps, prob = T, ylab = "f(x)", xlab = "x", main = "Histogram of draws from g(x) ")

lines(x, f(x), lty = 2)

```
2) Regression and Empirical Size

Regression 1,2: Read in the grocery retailer dataset.Name the dataset grocery. Use the least squares equationˆβ= (XTX)−1XTYto estimate regression model (2).To estimate the model, use the linear model function inR, i.e., uselm.
```{r}
grocery <- read.table("grocery.txt", header = TRUE, as.is = TRUE)
head(grocery)
n<-nrow(grocery)
a=ncol(grocery)

X <- cbind(rep(1,52), grocery$X1, grocery$X2, grocery$X3)
beta_hat <- solve((t(X) %*% X)) %*% t(X) %*% grocery$Y
round(t(beta_hat), 2)

lm0 <- lm(Y ~ X1 + X2 + X3, data = grocery)
lm0

```
Regression 3: Use R to estimate σ2, i.e., computeMSE=1n−4∑ni=1(Yi−ˆYi)2.  To perform this task,use theresidualsfunction.
```{r}
mse<- mean(lm0$residuals^2)
mse
#summary(lm0)

```
2.2 Test Slope

```{r}
summary(lm0)
```

2.3 Sampling Distribution

1)
```{r}
R<- 10000
stat.list<-rep(NA,R)
beta.list<-rep(NA,R)
for (i in 1:R){
  y.sim<- 4200-0*grocery$X1+15*grocery$X2+620*grocery$X3 + rnorm(52, mean=0, sd = sqrt(20500))
  model.sim<-lm(y.sim~grocery$X1+grocery$X2+grocery$X3)
  stat.list[i]<-summary(model.sim)[[4]][2,3]
  
}
#stat.list
head(stat.list)


```
2)
```{r}
hist(stat.list, probability = TRUE, breaks = 40)
t.list2<-seq(-5,5,by=.01)
lines(t.list2, dt(t.list2, 48), col = "green")


quantile(stat.list, .9) #simulated quantile
qt(.9, 48) #true quantile

quantile(stat.list, .95) #simulated quantile
qt(.95, 48) #true quantile

quantile(stat.list, .99) #simulated quantile
qt(.99, 48) #true quantile
```


  

