---
title: "Review"
author: "Sharleen Price spp2122"
date: "2/28/2018"
output: pdf_document
---
```{r}
#import a dataset

#housing <- read.csv("NYChousing.csv", as.is=TRUE)
#<-readLine("honocode.txt")
titanic <-read.table("Titanic.txt", header = TRUE, as.is = TRUE)

```

```{r}
#getting na values
apply(is.na(titanic), 2, sum)

#clean the data set
titanic <- na.omit(titanic)

w <- c(-3, 20, 9, 2)
w[w > 3] #returns 20,9

```
The call substr(string, start = , stop = ) returns the substring
from character position start to stop the given string.
substr(phrase, start = 8, stop = 12)

grep("students", HC) returns index of pattern
head(grepl("students", HC), 15) returns true or false vector based on pattern

for (i in 1:length(name_split)){
  split_comma<-append(split_comma,name_split[[i]][2])
}
M[M[,3] < 8, ]  data[truthVector,]

mean(t, na.rm = TRUE)
ifelse()
Versicolor <- ifelse(iris[,5]=="versicolor",1,0), if true->1,else->0

Switzerland.strikes <- subset(strikes, country == "Switzerland")


```{r}
#construct a base r plot

#lab 3 question 3
boxplot(Sepal.Width~Versicolor,main="Sepal of Versicolor vs Others",ylab="width",data=iris)
#boxplot(y~x, )

```
plot(Switzerland.strikes$year,Switzerland.strikes$inflation,
      main="Switzerland's Inflation Rate vs Time", xlab="Time",ylab="Inflation Rate",
      col="blue",type="l")  
hist(normal100)

plot(xval,yval, color, xlab = "Log of Units", ylab = "Log of Value", main = "Title")
legend("bottomright", legend = levels(factor(housing$after1950)), fill = unique(factor(housing$after1950)))

plot(BK$logUnits, BK$logValue, xlab="log(Units)", ylab="log(Value)", main="log(Property Value) vs log(Property Units)", ylim=c(9,21),col=11)
points(Manhattan$logUnits, Manhattan$logValue, col=4)
legend("bottomright", legend = c("Brooklyn","Manhattan"), fill=c(11,4))

```{r}
#construct a ggplot
library(ggplot2)

```
***Lab 5***
gplot(data = iris)+
geom_point(mapping = aes(x=Sepal.Length,y=Petal.Length,color = Species))+
  geom_abline(data = my.slopes,mapping=aes(slope=Sepal.Length,intercept=X.Intercept.,color=Species) )+ labs(title = "Sharleen's Plot") +  geom_point(aes(x=4.9,y=4.5), shape = 8, color =4) +geom_text(x = 4.9, y = 4.8, label = "(4.9, 4.5)", color = 4)+geom_point(aes(x=5.1,y=3), shape = 8, color =3) +geom_text(x = 5.1, y = 2.7, label = "(5.1, 3)", color = 3)+geom_point(aes(x=5.1,y=3), shape = 8, color =3) +geom_point(aes(x=5.8,y=1.2), shape = 8, color =2)+geom_text(x = 5.8, y = 1, label = "(5.8, 1.2)", color = 2)
  
  ggplot(data = wtid)+
geom_line(mapping = aes(x=Year,y=P99,color = "P99"))+geom_line(mapping = aes(x=Year,y=P99.5, color = "P99.5"))+geom_line(mapping = aes(x=Year,y=P99.9, color = "P99.9")) + labs(x = "Time" ,y="Income Treshold", title="Plot")

ggplot(data=titanic) +
  geom_bar(aes(x=Sex,fill=factor(Survived)))+
  labs(title = "Title",fill="Survived")
  
  ggplot(data=titanic) +
  geom_bar(aes(x=factor(Survived),fill=factor(Survived)))+
  facet_grid(~Sex)+
  labs(title = "Title",fill="Survived",x="")
  
 
```{r}
#use charater string functions

```
 
take a loop and write as one or two lines of code
```{r} 
#homework 1 question v
classes2 <- sort(unique(titanic$Pclass))
Pclass.Survival2 <- vector("numeric", length = 3)
names(Pclass.Survival2) <- classes

Pclass.Survival2<-round((tapply(titanic$Survived,titanic$Pclass,mean)),2)


```
 
*** Homework 2 question v***
 manhat.props <- c()
    for (props in 1:nrow(housing)) {
      if (housing$Borough[props] == "Manhattan") {
        manhat.props <- c(manhat.props, props)
      }
}
    med.value <- c()
    for (props in manhat.props) {
      med.value <- c(med.value, housing$Value[props])
    }
    med.value <- median(med.value, na.rm = TRUE)
    
    med.value
    
    median(housing$Value[housing$Borough=="Manhattan"],na.rm = TRUE) 
    
classes2 <- sort(unique(titanic$Pclass))
Pclass.Survival2 <- vector("numeric", length = 3)
names(Pclass.Survival2) <- classes

Pclass.Survival2<-round((tapply(titanic$Survived,titanic$Pclass,mean)),2)


***lab 3 question 6***
for (j in measurements) {
  MeanFlowers[j,]<-tapply(iris[,j],iris$Species,mean)
}
MeanFlowers


```{r}
?sapply
```

first.element <- function(vec){
  return(vec[1])
}

?sapply()

***Giuli's lab 6 Question 3***
titanic$Title <- (sapply(strsplit(titanic$Name,split=", "),FUN=function(vec) vec[2]))
titanic$Title <- (sapply(strsplit(titanic$Title,split=" "),FUN=function(vec) vec[1]))

titles <- c("Miss.","Mrs.","Mr.","Master.")
other_title_logical <- !(titanic$Title %in% titles)
titanic$Title[other_title_logical] <- "Other."

head(titanic$Title,10)

 